---
title: "Energy Efficiency data analysis"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

The aim is to perform energy analysis using 12 different building shapes
simulated in Ecotect in order to get efficient building designs.

The dataset contains eight attributes (or features, denoted by X1...X8) and two
responses (or outcomes, denoted by y1 and y2). 

Specifically:
X1 Relative Compactness: volume to surface ratio $Rc = 6 * V^{2/3} * S^{-1}$ 
V: volume (constant and equal to 771.75)
S:the Surface Area

X2: Surface Area
X3: Wall Area
X4: Roof Area
X5: Overall Height
X6: Orientation: Factor with 4 levels "North", "Easth", "South", "West"
X7: Glazing Area: Factor with 4 levels 0%, 10%, 25%, 40%
X8: Glazing Area Distribution: factor with 6 levels "None", "Uniform", "North", "Easth", "South", "West"
y1: Heating Load
y2: Cooling Load

```{r}
library(readxl)
ENB2012_data <- read_excel("ENB2012_data.xlsx")
ENB2012_data <- as.data.frame(ENB2012_data)
colnames(ENB2012_data) <- c("Relative_compactness", "Surface_Area", "Wall_Area",
                         "Roof_Area", "Overall_Height", "Orientation",
                         "Glazing_Area", "Glazing_Area_Distribution", 
                         "Heating_Load", "Cooling_Load")


ENB2012_data$Orientation <- as.factor(ENB2012_data$Orientation)
levels(ENB2012_data$Orientation) <- c("North", "Easth", "South", "West")
ENB2012_data$Glazing_Area <- as.factor(ENB2012_data$Glazing_Area)
levels(ENB2012_data$Glazing_Area) <- c("0%", "10%", "25%", "40%")
ENB2012_data$Glazing_Area_Distribution <- as.factor(ENB2012_data$Glazing_Area_Distribution)
levels(ENB2012_data$Glazing_Area_Distribution) <- c("None", "Uniform", "North", "Easth", "South", "West")
ENB2012_data$Glazing_Area_Distribution[ENB2012_data$Glazing_Area_Distribution=="0"] <- "None"
ENB2012_data$Glazing_Area_Distribution[ENB2012_data$Glazing_Area_Distribution=="1"] <- "Uniform"
ENB2012_data$Glazing_Area_Distribution[ENB2012_data$Glazing_Area_Distribution=="2"] <- "North"
ENB2012_data$Glazing_Area_Distribution[ENB2012_data$Glazing_Area_Distribution=="3"] <- "Easth"
ENB2012_data$Glazing_Area_Distribution[ENB2012_data$Glazing_Area_Distribution=="4"] <- "South"
ENB2012_data$Glazing_Area_Distribution[ENB2012_data$Glazing_Area_Distribution=="5"] <- "West"

head(ENB2012_data)

```
```{r}
summary(ENB2012_data)
```
**RESPONSES**
```{r}
library(ggplot2)
ggplot(ENB2012_data, aes(x = Heating_Load))+
  geom_histogram(binwidth = 3, fill = "#404040", color = "black", alpha = 0.9)+
  theme_bw()+
  labs(x = NULL,
       y = NULL,
       title = "Histogram of Heating Load")

ggplot(ENB2012_data, aes(x = Cooling_Load))+
  geom_histogram(binwidth = 3, fill = "#004C99", color = "black", alpha = 0.9)+
  theme_bw()+
  labs(x = NULL,
       y = NULL,
       title = "Histogram of Cooling Load")
```
**CORRELATION**
```{r}
corr <- cor(ENB2012_data[,-c(6,7,8)], method = "spearman"); corr
```
```{r}
library(ggcorrplot)
ggcorrplot(corr, hc.order = TRUE, type = "lower", lab = TRUE)
```

Relative Compactness and Surface Area are perfectly inversely proportional, 
which is because in the simulations we have assumed that the volume of the 
buildings is constant (there is an analytic formula linking the surface area to 
RC and volume). 

From the data we also have:

$$Surface Area = Wall Area + 2 * Roof Area$$ 

Therefore, we have a linear combination for these 3 variablese (Not detected by the 
correlation matrix because it checks only pairwise)

Roof Area is also highly correlated with Overall Height: they share a correlation
coefficient of -0.93 indicating high negative relationship: this is due to the 
fact that the different buildings shapes have all the same volume.
Hence, buildings with a large roof area can't be high and viceversa. 


We can also notice that the two responses Cooling Load and Heating Load have an high 
level of correlation: hence we can state that the effect of the predictors on both 
the responses is almost the same. 

```{r}
cor(ENB2012_data$Heating_Load, ENB2012_data$Cooling_Load, method = "spearman")
```
**VIF ANALYSIS**

Let's perform the vif analysis by just removing the perfect correlated predictors 
Surface Area and Roof Area.
```{r}
z <- lm(Heating_Load~Relative_compactness+Overall_Height+Orientation+
          Glazing_Area+Glazing_Area_Distribution+Wall_Area, data = ENB2012_data)

library(faraway)
faraway::vif(z)
```
We can still see that there's a problem of collinearity among Relative Compactness,
Overall Height and Wall Area.
Let's now try to remove Wall Area
```{r}
a <- lm(Heating_Load~Relative_compactness+Overall_Height+Orientation+
          Glazing_Area+Glazing_Area_Distribution, data = ENB2012_data)
faraway::vif(a)

```

At this stage Relative compactnessa and Overall Height looks fine.
Consider now the two factor variables Glazing Area and Glazing Area Distribution 
where there seems to be correlation. 
As said before Glazing Area refers to the percentage of glazing the building has 
while Glazing Area Distribution is related to its distribution. It seems natural
to state that for buildings with 0% glazing area, there will not be any distribution.


**GLAZING AREA DISTRIBUTION**
```{r}
par(mfrow=c(1,2))
boxplot(Heating_Load ~ Glazing_Area_Distribution, data = ENB2012_data)
boxplot(Cooling_Load ~ Glazing_Area_Distribution, data = ENB2012_data)

ggplot(ENB2012_data, aes(Glazing_Area_Distribution, Heating_Load))+
  geom_boxplot()+
  geom_point(aes(colour = Glazing_Area))+
  theme_bw()
```

```{r}
b<- lm(Heating_Load ~ Glazing_Area_Distribution, data = ENB2012_data)
summary(b)
```
```{r}
ENB2012_data$Glazing_Area_Distribution <- relevel(factor(ENB2012_data$Glazing_Area_Distribution), ref = 2)
c <- lm(Heating_Load ~ Glazing_Area_Distribution, data = ENB2012_data)
summary(c)
```
No significant difference of Heating Load with respect to the different levels of Orientation
is displayed: the only significant difference lays between having a distribution, 
namely having glazing, and not having a distribution, so not having glazing. 
The p-value of the F-test is indeed 3.69e-06: therefore, when taken alone, the
predictor seems to be significant and to affect the responses.

**GLAZING AREA**
Let's now investigate deeper the variable Glazing Area
```{r}
d <- lm(Heating_Load ~ Glazing_Area, data = ENB2012_data)
summary(d) 
```


```{r}
boxplot(Heating_Load ~ Glazing_Area, data = ENB2012_data, xlab = "Glazing Area",
        ylab = "Heating Load")

ggplot(ENB2012_data, aes(Glazing_Area, Heating_Load))+
  geom_boxplot()+
  geom_point(aes(colour=Glazing_Area_Distribution))+
  theme_bw()
```
The percentage of glazing area on the buildings seems to affect the response Heating Load. 
P-value of the F-test is indeed 5.09e-10, which leads us to reject the null hypotheses
that the there is no significant difference between the null model and the model with the
predictor.

We now want to check if the two predictors, Glazing Area and Glazing Area
Distribution, share some kind of interaction between each other in order to create a proper
model taking into account both of them.  
In this way we are testing if the factor Glazing Area Distribution is still
significant and must be considered for our regression model.
```{r}
interaction.plot(ENB2012_data$Glazing_Area_Distribution, ENB2012_data$Glazing_Area,
                 ENB2012_data$Heating_Load, xlab = "Glazing Area Distribution",
                 ylab = "Heating Load", col = c(1:4), legend = T, trace.label = "Glazing Area" )
```
Since lines are parallel there does not seem to be any interaction. 

```{r}
e <- aov(Heating_Load~Glazing_Area*Glazing_Area_Distribution, ENB2012_data)
summary(e)
```
Implementing a model with the interaction term shows that it is not signficant. 

```{r}
e <- lm(Heating_Load~Glazing_Area + Glazing_Area_Distribution + Orientation, data = ENB2012_data)
anova(e)
```
We can observe that the factor Glazing Area Distribution is still not significant
and hence can be removed from the model since it can be resumed from the variable
Glazing Area.


**MODEL SELECTION**
```{r}
library(leaps)
regs <- regsubsets(Heating_Load~Relative_compactness+Overall_Height+
          Glazing_Area+Orientation, data = ENB2012_data)
rs <- summary(regs)
rs$which
```

```{r}
n <- nrow(ENB2012_data)
AIC <- n*log(rs$rss/n) + (1:8)*2
plot(1:8, AIC, ylab = "AIC", xlab = "Number of parameters")
```
Since we want to minimise the AIC, the above suggest to take a model with 5 predictors,
that would be Relative Compactness, Overall Height and Glazing Area. 
With regards to Glazing Area, within the regsubsets analysis each level is considered to be
a single predictor, which are four but one, the reference level, stays with the intercept. 

```{r}
plot(1:8, rs$bic, ylab = "BIC", xlab = "Number of parameters")
```
Even with the BIC analysis, we end up with the same conclusion that Orientation is not 
worthwhile to be included in the model. 


**ORIENTATION**

```{r}
par(mfrow= c(1,2))
boxplot(Heating_Load ~ Orientation, data = ENB2012_data, ylab = "Heating Load")
boxplot(Cooling_Load ~ Orientation, data = ENB2012_data, ylab = "Cooling Load")


ggplot(ENB2012_data, aes(Orientation, Heating_Load))+
  geom_boxplot()+
  geom_point()+
  theme_bw()
```

```{r}
f <- lm(Heating_Load ~ Orientation, data = ENB2012_data)
summary(f)
```
Indeed, there does not seem to be any difference within the four levels of Orientation
since the boxplots overlap and the p-value of the F test is almost 1, leading to the 
conclusion that there is no significant difference between the null model and the model with
the predictor Orientation. 


**BROKEN STICK REGRESSION FOR RELATIVE COMPACTNESS**

```{r}
plot(Heating_Load ~ Relative_compactness, data = ENB2012_data)
```
We can observe 2 different patterns for buildings with small and large RC: in 
particular a decreasing relationship for RC from 0.6 to 0.75 and then a jump, 
following again a decreasing pattern. 
```{r}
g <- lm(Heating_Load~Relative_compactness, data = ENB2012_data)
plot(Heating_Load~Relative_compactness, data = ENB2012_data)
abline(g, col = 2, lwd = 2)
```
With regards to the overall effect of RC on HL we can although observe a positive
relationship between HL and RC, due to the jump we can observe at RC ~ 0.75.


Let's now try to separate the patterns considering two different subsets.
```{r}
h <- lm(Heating_Load ~ Relative_compactness , data = ENB2012_data, subset = Relative_compactness < 0.75)
summary(h)

```

```{r}
i <- lm(Heating_Load ~ Relative_compactness, data = ENB2012_data, subset = Relative_compactness >= 0.75)
summary(i)
```

```{r}
beta1 <- coef(h); beta2 <- coef(i)
plot(Heating_Load ~ Relative_compactness, data = ENB2012_data,
     xlab = "Relative Compactness", ylab = "Heating Load")
abline(v=0.75)
segments(0.6,beta1[1]+beta1[2]*0.6,0.75,beta1[1]+beta1[2]*0.75, col = "blue", lwd = 2)
segments(0.75,beta2[1]+beta2[2]*0.75,1,beta2[1]+beta2[2]*1, col = "blue", lwd = 2)

B_l <- function(x) ifelse(x<0.75, 0.75-x, 0)
B_r <- function(x) ifelse(x>0.75, x-0.75, 0)
lmod <- lm(Heating_Load~B_l(Relative_compactness)+B_r(Relative_compactness),
           data = ENB2012_data)
summary(lmod)
x <- seq(0.5:1, by = 0.01)
beta <- coef(lmod)
py <- beta[1]+beta[2]*B_l(x)+beta[3]*B_r(x)
lines(x,py,col=2, lwd = 2)
```
When dealing with a broken stick regression imposing the continuity 
changes the sign of the slope for both RC patterns.

From a deeper analysis we can observe that the leap we see at point
RC=0.75 is due to the fact that all buildings with RC<0.75 have a value of Overall
Height of 3 while all buildings with RC>0.75 have Overall Height = 7:

Let's then try to perform a broken stick regression without imposing the continuity
but only separating the 2 patterns.
```{r}
B2_l <- function(x) ifelse(x<0.75, x, 0)
B2_r <- function(x) ifelse(x>0.75, x, 0)
lmod_nc <- lm(Heating_Load~B2_l(Relative_compactness)+B2_r(Relative_compactness),
           data = ENB2012_data)
summary(lmod_nc)
beta3 <- coef(lmod_nc)
plot(Heating_Load ~ Relative_compactness, data = ENB2012_data)
abline(v=0.75)
segments(0.5,beta3[1]+beta3[2]*0.5,0.75,beta3[1]+beta3[2]*0.75, col = "blue", lwd = 2)
segments(0.75,beta3[1]+beta3[3]*0.75,1,beta3[1]+beta3[3]*1, col = "blue", lwd = 2)
```

**OVERALL HEIGHT**

```{r}
plot(Heating_Load~Overall_Height, data = ENB2012_data)
plot(Overall_Height~Relative_compactness, data = ENB2012_data)
abline(v=0.75, col = 2)
```
We can indeed see that Overall Height only takes 2 values, 3.5 and 7.0;
it hence represents buildings with 1 or 2 floors, namely buildings with RC<0.75 and
RC>0.75. 
In other words,considering then a broken stick regression to separate the
behaviours for RC smaller and greater than 0.75 means separating buildings with 
Overall Height of 3.5 and 7. 
For this reason, it does not make sense anymore to consider both
the variable Overall Height and the broken stick regression for Relative Compactness as
they explain the exact same thing and we would have unidentifiability problems. 
```{r}
lm <- lm(Heating_Load~B2_l(Relative_compactness)+B2_r(Relative_compactness)+
           Glazing_Area+Overall_Height,
           data = ENB2012_data)
vif(lm)
```
Indeed VIFs literally blow up. 





We now have two ways to proceed: consider Overall Height as continuous predictor
or consider it as factor with 2 levels since we con only observe 2 values (3.5 and 7). 
This is due to the fact that builidngs are generated by simulation using Ecotect: 
since we would like to study the general effect of the input variables to determine
the output variables such as Heating Load and Cooling Load, we choose to keep 
Overall Height as continuous variable not to overfit the model and not to reduce 
the study only to buildings with 1 or 2 floors as in real life those could be higher. 


```{r}
mod <- lm(Heating_Load~Relative_compactness+Glazing_Area+Overall_Height,
              data = ENB2012_data)
summary(mod)
```

**DIAGNOSTIC**
```{r}
plot(mod)
```
**HETEROSCEDASTICITY**
We can see that variance is increasing as fitted values get bigger. 
Let's then try to stabilise the variance by transforming our responses. 

Since we have strictly positive responses we can look for a Box-Cox transformation.

```{r}
library(faraway); library(MASS)
boxcox(mod, plotit = T)
boxcox(mod, plotit = T, lambda = seq(-0.3, 0.3, by = 0.1))

```
The confidence intervals associated with lambda does not include 1 but rather 0:
this suggests us to apply a logarithmic transformation. 
```{r}
logmod <- lm(log(Heating_Load)~0+Relative_compactness+Glazing_Area+Overall_Height,
         data = ENB2012_data)
plot(logmod)
summary(logmod)
CLlogmod <- lm(log(Cooling_Load)~0+Relative_compactness+Glazing_Area+Overall_Height,
         data = ENB2012_data)
summary(CLlogmod)
```
The situation has improved but we can still observe slight 
heteroscedasticity

**INDIPENDENCE AND IDENTICAL DISTRIBUTION OF ERROR**
Since we are dealing with a cross-sectional data, namely we are comparing the 
differences among selected objects (the buildings) with no regards to differences
in time there's no need to look for auto-correlation. 

Assuming hence that errors are independent but not identically distributed 
(allowing different variances), we can use weighted least squares (WLS). 
We do not know a priori the weights: we then try using as weights the reciprocal
of the square of the estimates of the error standard deviations.

```{r}
res <- logmod$residuals; fit <- logmod$fitted.values

sd.func <- lm(abs(res)~fit)
sd <- sd.func$fitted.values #estimates of the error standard deviations
weights <- 1/sd^2
wls_mod <- lm(log(Heating_Load)~0+Relative_compactness+Glazing_Area+Overall_Height,
             data = ENB2012_data,
             weights = weights)
plot(wls_mod)
summary(wls_mod)
```
We observe that the situation has slightly improved but we have not reached the 
perfect homoscedasticity.
Indeed the Breusch-Pagan Test still rejects the null hypotheses that residuals 
are homoscedastic.
```{r}
library(lmtest)
bptest(wls_mod)
```
Another possible reason for the heteroscedasticity could be that predictors and 
responses could not be linearly related. 
Let's then make the partial residual plot, which display the residuals of each
predictor against the response variable. 
The blue line shows the expected residuals if the relationship between the predictor
and response variable was linear. The pink line shows the actual residuals.

If the two lines are significantly different, then this is evidence of a nonlinear 
relationship.

```{r}
library(car)
crPlots(logmod)
```
We can see that residuals for both Relative Compactness and Overall Height are
linear. 

**NORMALITY**
```{r}
qqnorm(residuals(wls_mod))
qqline(residuals(wls_mod))
shapiro.test(residuals(wls_mod))
```
We can observe that errors are not normally distributed but this is not a problem 
since we have a large number of observation and hence the issue can be handles 
with the Central Limit Theorem.


**LEVERAGES**
We can sort the values and look for the values that are greater than 2p/n = 14/768. 
They are faraway from the remaining buildings.

```{r}
hatv <- hatvalues(logmod)
buildings <- rownames(ENB2012_data)
halfnorm(hatv, labs = buildings, ylab = "Leverages")
abline(h = 2*sum(hatv)/nrow(ENB2012_data))
which(hatv>2*sum(hatv)/nrow(ENB2012_data))
```


**OUTLIERS**
```{r}
cr_value <-  qt(0.05/(2*768), 768, lower.tail = F)
sr <- rstudent(logmod)
which(abs(sr)>cr_value)
```
No outlier is detected.  

*InFLUENTIAL OBSERVATIONS*
```{r}
cook <- cooks.distance(logmod)
halfnorm(cook,4, labs = buildings, ylab = "Cook's distances" )

```
Let's now see if having or not having these buildings will produce a big change in our model.

Excluding buildings and comparing outputs

```{r}
logmodi <- lm(log(Heating_Load)~ 0 + Relative_compactness+Glazing_Area+
             Overall_Height,
         data = ENB2012_data, subset = (buildings != c("19","18","20", "17")))
sumary(logmod)
sumary(logmodi)
```
Not influential at all.


**EXTERNAL VALIDATION**
We are now focused on the predictive power of our model, namely how good it will
be at predicting new observations.
Since we do not have a new dataset we split randomly our observation into two halves
and fit the model on the training dataset.
```{r}
library(boot)
set.seed(1)
train <- sample(nrow(ENB2012_data), nrow(ENB2012_data)/2)
test <- ENB2012_data[-train,]
pred_mod <- lm(log(Heating_Load)~0+Relative_compactness+Glazing_Area+Overall_Height,
         data = ENB2012_data, subset = train)
pred <- predict(pred_mod, newdata = test)
plot(pred, test$Heating_Load)
MSE <- mean((test$Heating_Load-pred)^2)
glm.fit <- glm(Heating_Load~Relative_compactness+Glazing_Area+Overall_Height,
               data = ENB2012_data)
mse <- cv.glm(ENB2012_data, glm.fit)
mse$delta
```







































